Enforced Invariant Confluence

### 2018-04-25 doc/paper/sections/evaluation.tex
Updated benchmarks.

\example[$\ints^2$]\examplelabel{TwoIntsEval}
We begin with a minimal working example. Consider again our recurring example
of $\ints^2$ from \exampleref{Z2}. The Python code used to describe the object,
transactions, and invariant is given in \figref{Z2Code}. When we call
\python{checker.check()}, the interactive decision procedure produces a
counterexample in less than a tenth of a second.  After we label the
counterexample and refine the invariant with $y \leq 0$, the interactive
decision procedure determines that the object is \invariantconfluent{}, again,
in less than a tenth of a second. Note that the object is \invariantconfluent{}
but \emph{not} \invariantclosed{}, so prior work that relies on
\invariantclosure{} alone to determine \invariantconfluence{} would not be able
to identify this example as \invariantconfluent{}.

@@

\subsection{Segmented Invariant Confluence}%
\seclabel{SegmentedInvariantConfluenceEval}
Now, we evaluate the performance of replicating an object with segmented
\invariantconfluence{} as compared to replicating it with eventual consistency
or linearizability.

\begin{figure}[ht]
  \centering

  \begin{subfigure}[c]{\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{figures/vary_withdraws.pdf}
  \end{subfigure}
  \begin{subfigure}[c]{\columnwidth}
    \includegraphics[width=\columnwidth]{figures/vary_segments.pdf}
  \end{subfigure}

  \caption{%
    Segmented \invariantconfluent{} replication throughput versus coordination,
    induced by executing disallowed transactions (top) and by transitioning
    across segments (bottom).
  }\figlabel{ThroughputVsGlobalSyncs}
\end{figure}

\begin{benchmark}\benchlabel{VaryWithdraws}
  Consider again the PN-Counter from \exampleref{CounreachableExample} and the
  corresponding transactions, invariants, and single-segment segmentation that
  forbids concurrent decrements. We replicate this object on 32 servers
  deployed on 32 m5.xlarge EC2 instances within the same availability zone.
  Each server has three colocated clients that issue increment and decrement
  transactions. We replicate the object with eventual consistency, segmented
  \invariantconfluence{}, and linearizability and measure the system's total
  throughput as we vary the fraction of client requests that are decrements.
  The results are shown in the top of \figref{ThroughputVsGlobalSyncs}.

  Both eventually consistent replication and linearizable replication are
  unaffected by the workload, achieving roughly 700,000 and 7,000 transactions
  per second respectively. Segmented \invariantconfluent{} replication performs
  well for low-decrement workloads and performs increasingly poorly as we
  increase the fraction of decrement transactions, eventually performing worse
  than linearizable replication. For example, with 5\% decrement transactions,
  segmented \invariantconfluent{} replication performs an order of magnitude
  better than linearizable replication; with 50\% decrements, it performs as
  well; and with 100\% decrements, it performs two times worse.

  These results are expected. Increment transactions can execute without any
  coordination while decrement transactions require global coordination. As we
  increase the fraction of decrements, we increase the amount of coordination
  that the system has to perform which in turn drastically decreases the
  throughput. These results also offer two insights.
  %
  First, for low-decrement workloads, segmented \invariantconfluent{}
  replication achieves a compromise between strong and weak consistency. It
  guarantees that invariants are maintained (which is impossible with eventual
  consistency if the object is not \invariantconfluent{}) with performance many
  times better than strongly consistent replication.
  %
  Second, segmented \invariantconfluent{} replication is poorly suited to
  workloads that require a large amount of coordination. For workloads without
  much inherit concurrency (e.g.\ decrement-mostly workloads), maintaining
  invariants is best done with strong consistency. It provides stronger
  guarantees with better performance.
\end{benchmark}

\begin{benchmark}\benchlabel{VarySegmentLength}
  Consider again the object, transactions, and invariants from \exampleref{Z2}
  and \exampleref{SegmentedZ2}. As with \benchref{VaryWithdraws}, we replicate
  the object across 32 servers. Clients issue 50\% increment $x$ transactions,
  and 50\% decrement $y$ transactions. We consider a ``checkerboard''
  segmentation $\Sigma_n = \setst{(I_{i, j}, T)}{i, j \in \ints}$ where segment
  invariant $I_{i, j}$ consists of the square of points $\setst{(x, y)}{ni \leq
  x < n(i + 1), nj \leq y < n(j + 1)}$ with side length $n$. For example,
  $\Sigma_1$ places each point in its own segment, $\Sigma_2$ tessellates
  $\ints^2$ with 2x2 squares, $\Sigma_3$ tessellates $\ints^2$ with 3x3
  squares, and so on. We measure the throughput of the object replicated with
  eventually consistent, segmented \invariantconfluent{}, and linearizable
  replication as we vary the segment side length $n$. The results are shown in
  the bottom \figref{ThroughputVsGlobalSyncs}.

  This benchmark tells the same tale as \benchref{VaryWithdraws}. Eventual
  consistency and linearizability are unaffected by workload, and eventual
  consistency outperforms linearizability by roughly two orders of magnitude.
  In this example, the segmented \invariantconfluent{} replication only
  requires coordination when transitioning between segment boundaries, so as we
  increase the segment side length, the throughput of the system increases
  significantly.
\end{benchmark}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\columnwidth]{figures/vary_nodes.pdf}
  \caption{%
    Throughput of eventually consistent, segmented \invariantconfluent{}, and
    linearizable replication measured against the number of
    nodes.
  }\figlabel{VaryNodes}
\end{figure}

\begin{benchmark}
  In this benchmark, we measure the scale-out of segmented
  \invariantconfluent{} replication. We repeat \benchref{VaryWithdraws} with a
  10\% decrement rate, but this time we vary the number of servers we use to
  replicate our object.  When we replicate with $n$ servers, we use $3n$
  clients (the $3$ colocated clients on each server) as part of the workload.
  The results are shown in \figref{VaryNodes}.

  Eventually consistent replication scales perfectly with the number of nodes,
  confirming the results in~\cite{bailis2014coordination}.  Linearizable
  replication, on the other hand, scales up to about 3--5 servers before
  performance begins to decrease. These numbers are consistent with typical
  deployments of state-machine replication protocols like
  Paxos~\cite{chandra2007paxos}. Segmented \invariantconfluent{} replication
  scales up to about 6--8 servers before succumbing to the same scalability
  bottlenecks as linearizable replication.

  These results highlight the importance of coordination avoidance in
  distributed databases. While segmented \invariantconfluent{} replication
  scales out slightly better than linearizable replication, both scale
  significantly worse than eventually consistent replication even for a very
  low (i.e.\ 10\%) decrement workload. This demonstrates that even a small
  amount of coordination can significantly reduce the scalability of a system.
\end{benchmark}

>>>>

\example[$\ints^2$]\examplelabel{TwoIntsEval}
We begin with a minimal working example. Consider again our recurring example
of $\ints^2$ from \exampleref{Z2}. The Python code used to describe the object,
transactions, and invariant is given in \figref{Z2Code}. When we call
\python{checker.check()}, the interactive decision procedure produces a
counterexample $s_1 = (0, 1), s_2 = (1, 0)$ in less than a tenth of a second
and automatically recognizes that $s_2$ is reachable. After we label $s_1$ as
unreachable and refine the invariant with $y \leq 0$, the interactive decision
procedure determines that the object is \invariantconfluent{}, again, in less
than a tenth of a second. Note that the object is \invariantconfluent{} but
\emph{not} \invariantclosed{}, so prior work~\cite{li2012making,
li2014automating, balegas2015towards, gotsman2016cause} that relies on
\invariantclosure{}---or another equivalent sufficient condition---to determine
\invariantconfluence{} would not be able to identify this example as
\invariantconfluent{}.

@@

\subsection{Segmented Invariant Confluence}%
\seclabel{SegmentedInvariantConfluenceEval}
Now, we evaluate the performance of replicating an object with segmented
\invariantconfluence{} as compared to the performance of replicating it with
eventual consistency or linearizability.
%
There are two hypotheses about the performance of segmented
\invariantconfluent{} replication that we aim to confirm.
%
First, segmented \invariantconfluent{} replication provides higher throughput
and better scalability than linearizabile replication for workloads which
require little coordination (i.e.\ low-coordination workloads).
%
Second, the throughput and scalability of segmented \invariantconfluent{}
replication decreases as we increase the fraction of transactions which require
coordination.

These hypotheses state that segmented \invariantconfluent{} is more performant
than linearizabile replication for low-coordination workloads. But by how much?
We also aim to measure the absolute performance and scalability benefits of
segmented \invariantconfluent{} replication and how they vary as we vary the
coordination required by a workload. We perform two controlled microbenchmarks
to confirm our hypotheses and discover the absolute performance benefits. The
workloads themselves are trivial, but also not salient. Our objective is to
obtain a controlled measure of throughput and scalability as we vary workload
contention.

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{figures/throughput_vs_fraction_16.pdf}
  \caption{%
    Segmented \invariantconfluent{} replication throughput versus coordination
    induced by executing disallowed decrement transactions.
  }\figlabel{ThroughputVsGlobalSyncs}
\end{figure}

\begin{benchmark}\benchlabel{VaryWithdraws}
  Consider again the PN-Counter from \exampleref{CounreachableExample} and the
  corresponding transactions, invariants, and single-segment segmentation that
  forbids concurrent decrements. We replicate this object on 16 servers
  deployed on 16 m5.xlarge EC2 instances within the same availability zone.
  Each server has three colocated clients that issue increment and decrement
  transactions. We replicate the object with eventual consistency, segmented
  \invariantconfluence{}, and linearizability and measure the system's total
  throughput as we vary the fraction of client requests that are decrements.
  The results are shown in \figref{ThroughputVsGlobalSyncs}.

  Both eventually consistent replication and linearizable replication are
  unaffected by the workload, achieving roughly 375,000 and 12,000 transactions
  per second respectively. Segmented \invariantconfluent{} replication performs
  well for low-decrement (i.e.\ low-coordination) workloads and performs
  increasingly poorly as we increase the fraction of decrement transactions,
  eventually performing worse than linearizable replication. For example, with
  5\% decrement transactions, segmented \invariantconfluent{} replication
  performs over an order of magnitude better than linearizable replication;
  with 50\% decrements, it performs as well; and with 100\% decrements, it
  performs two times worse.

  % These results are expected. Increment transactions can execute without any
  % coordination while decrement transactions require global coordination. As we
  % increase the fraction of decrements, we increase the amount of coordination
  % that the system has to perform which in turn drastically decreases the
  % throughput. For high contention workloads, segmented \invariantconfluent{} replication performs worse than linearizable replication

  These results offer two insights.
  %
  First, the relationship between segmented \invariantconfluent{} and
  linearizable replication is analogous to the relationship between pessimistic
  and optimistic concurrency control protocols with coordination playing the
  role of contention. Linearizable replication pessimistically assumes that
  concurrently executing \emph{any} pair of transactions will lead to an
  invariant violation. Thus, clients send transactions directly to a leader to
  be linearized. Conversely, segmented \invariantconfluent{} replication
  optimistically attempts to perform every transaction locally and without
  coordination. A server only initiates a round of coordination if it is found
  to be necessary. As a consequence, segmented \invariantconfluent{}
  replication can offer substantial performance benefits over linearizable
  replication for low-coordination workloads, but is inferior for medium to
  high contention workloads.

  Second, throughput does not decrease linearly with the amount of
  coordination. Even infrequent coordination can drastically decrease
  throughput. Increasing the fraction of decrements from 0\% to 1\% decreases
  throughput by a factor of 2. Increasing again to 3\%, the throughput
  decreases by another factor of 2. With 90\% coordination-free transactions
  (i.e.\ 10\% decrements), we achieve only 10\% of the throughput of eventual
  consistency.

%  %
%  First, for low-decrement workloads, segmented \invariantconfluent{}
%  replication achieves a compromise between strong and weak consistency. It
%  guarantees that invariants are maintained (which is impossible with eventual
%  consistency if the object is not \invariantconfluent{}) with performance many
%  times better than strongly consistent replication.
%  %
%  Second, segmented \invariantconfluent{} replication is poorly suited to
%  workloads that require a large amount of coordination. For workloads without
%  much inherit concurrency (e.g.\ decrement-mostly workloads), maintaining
%  invariants is best done with strong consistency. It provides stronger
%  guarantees with better performance.
\end{benchmark}

% \begin{benchmark}\benchlabel{VarySegmentLength}
%   Consider again the object, transactions, and invariants from \exampleref{Z2}
%   and \exampleref{SegmentedZ2}. As with \benchref{VaryWithdraws}, we replicate
%   the object across 32 servers. Clients issue 50\% increment $x$ transactions,
%   and 50\% decrement $y$ transactions. We consider a ``checkerboard''
%   segmentation $\Sigma_n = \setst{(I_{i, j}, T)}{i, j \in \ints}$ where segment
%   invariant $I_{i, j}$ consists of the square of points $\setst{(x, y)}{ni \leq
%   x < n(i + 1), nj \leq y < n(j + 1)}$ with side length $n$. For example,
%   $\Sigma_1$ places each point in its own segment, $\Sigma_2$ tessellates
%   $\ints^2$ with 2x2 squares, $\Sigma_3$ tessellates $\ints^2$ with 3x3
%   squares, and so on. We measure the throughput of the object replicated with
%   eventually consistent, segmented \invariantconfluent{}, and linearizable
%   replication as we vary the segment side length $n$. The results are shown in
%   the bottom \figref{ThroughputVsGlobalSyncs}.
%
%   This benchmark tells the same tale as \benchref{VaryWithdraws}. Eventual
%   consistency and linearizability are unaffected by workload, and eventual
%   consistency outperforms linearizability by roughly two orders of magnitude.
%   In this example, the segmented \invariantconfluent{} replication only
%   requires coordination when transitioning between segment boundaries, so as we
%   increase the segment side length, the throughput of the system increases
%   significantly.
% \end{benchmark}

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{figures/throughput_vs_num_nodes_multi.pdf}
  \caption{%
    Throughput of eventually consistent, segmented \invariantconfluent{}, and
    linearizable replication measured against the number of nodes for workloads
    with varying fractions of decrement transactions. For example, the
    ``segmented (0.2)'' line measures the performance of segmented
    \invariantconfluent{} replication with 20\% decrement transactions.
    Eventually consistent replication and linearizable replication are not
    affected by workload.
  }\figlabel{VaryNodes}
\end{figure}

\begin{benchmark}
  In this benchmark, we measure the scale-out of segmented
  \invariantconfluent{} replication. We repeat \benchref{VaryWithdraws} while
  we vary the number of servers that we use to replicate our object. When we
  replicate with $n$ servers, we use $3n$ clients (the $3$ colocated clients on
  each server) as part of the workload. The results are shown in
  \figref{VaryNodes}.

  Eventually consistent replication scales perfectly with the number of nodes,
  confirming the results in~\cite{bailis2014coordination}. Linearizable
  replication, on the other hand, scales up to about 3 servers before
  performance begins to decrease. Segmented \invariantconfluent{} replication
  scales well for low-coordination workloads and poorly for high-coordination
  workloads. For 1\%, 5\%, 20\%, and 50\% decrement transactions, segmented
  \invariantconfluent{} replication scales up to 24, 12, 4, and 1 server
  respectively.

  These results echo the results of \benchref{VaryWithdraws}. For
  low-coordination workloads, segmented \invariantconfluent{} replication can
  offer almost an order of magnitude better scalability of linearizable
  replication, but coordination decreases scalability superlinearly. Even
  infrequent coordination can drastically reduce the scalability of segmented
  \invariantconfluent{} replication with segmented \invariantconfluent{}
  replication ultimately scaling worse than linearizable replication for
  high-coordination workloads.
\end{benchmark}
